---
title: "Comparing model predictions and observed data using modelcheck"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing the model prediction and observed data using modelcheck}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

This vignette describes how to use the `modelcheck` package to specify a comparative layout that organizes the visual representation of model predictions and observed data into a layout to enhance visual comparison. We developed `modelcheck` to support comparative layouts of three types: juxtaposition, superposition, and explicit encoding. For a more general introduction to `modelcheck` and demonstration of its use in a standard model check workflow, see `vignette("modelcheck")`.

# Setup

The following libraries are required to run this vignette:

```{r setup}
library(dplyr)
library(purrr)
library(modelcheck)
library(ggplot2)
library(ggdist)
library(cowplot)
library(rstan)
library(brms)
library(gganimate)

theme_set(theme_tidybayes() + panel_border())
```

These options help Stan run faster:

```{r}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```


# Model

We use the built-in model `mpg_model`, a `brmsfit` object fitted in Gaussian model family with checkable push-forward transformations `mu` and `sigma`, to demonstrate different visual representations that `modelcheck` defines to show the distributions of model predictions and observed data. 

```{r example_model}
mpg_model
```

# Juxtaposition

Juxtaposition puts the visual representations of model predictions and observed data side by side. This comparative layout presents each object (visual representation for model predictions and observed data) separately, but requires that the viewer scan the plots to relate them. We specify the juxtaposition layout for a model check by `mc_layout_juxtaposition()`.

```{r juxtaposition}
mpg_model %>%
  mcplot() +
  mc_layout_juxtaposition() +
  mc_gglayer(coord_flip())
```

`modelcheck` also supports a variant of juxtaposition, nested juxtaposition, which helps reduce the amount of scanning required by viewers by nesting the juxtaposition layout in the plot when the model check has a discrete conditional variable. We specify this by `mc_layout_nested()`.

```{r nested juxtaposition}
mpg_model %>%
  mcplot() +
  mc_condition_on(x = vars(vs)) +
  mc_model_auto(n_sample = 1) +
  mc_layout_nested()
```

# Superposition

Superposition overlays the objects (visual representations for model predictions and observed data) and presents the objects in a single coordinate system. `modelcheck` uses superposition as the default comparative layout if users don't specify a comparative layout.

```{r superposition}
mpg_model %>%
  mcplot() +
  mc_layout_superposition() +
  mc_gglayer(coord_flip())
```

# Explicit encoding

Explicit encoding directly encodes the connections between objects visually after calculating the difference between model predictions and observed data. For example, a residual plot presents the difference between the model predictions and observed data to encode the comparison about a central line encoding perfect prediction. A quantile-quantile (QQ) plot computes the quantiles of the residuals to encode the comparison. 

Explicit encoding can also explicitly enable the user to check a model feature but not depends on viewer's perception on the distribution of model prediction and observed data shown in raw coordinates of response variable and independent variables. For example, residual plot conditional on response variable enables to check linearity and QQ plot enables to check the normality of residuals.

Let's check the linearity of the model by specifying a residual plot conditional on the response variable `mpg`, where we use two visual representations (`mc_model_lineribbon()` and `mc_model_point()`) for the model predictions to reveal the trend of residuals while also showing a set of raw data points.

```{r residual}
mpg_model %>% 
  mcplot() + 
  mc_condition_on(x = vars(mpg)) + 
  mc_layout_encoding("residual") + 
  mc_model_lineribbon(alpha = .2) + 
  mc_model_point(n_sample = 1) + 
  mc_gglayer(geom_hline(yintercept = 0))
```

Next let's check the normality of the residuals by drawing a QQ plot. We could specify this easily by using `mc_layout_encoding(tranform = "qq")`.

```{r qq}
mpg_model %>% 
  mcplot() + 
  mc_layout_encoding("qq") + 
  mc_model_point(n_sample = 1) + 
  mc_gglayer(geom_abline())
```

## Specify your own explicit encoding

We could use `modelcheck` to specify a customized transformation function that computes the comparison between model predictions and observed data. The transformation function should take as input a data frame that is generated from the `newdata` data frame in `mc_distribution()` (if not specified, it's the data used to fit the model, e.g., you can get that by `insight::get_data(model)`) with additional columns: `.row`, `.draw`, `prediction`, and `observation` (if `newdata` has a column for the response variable). We can transform the input data frame to generate a new one that has the columns for the variable shown on the y-axis and x-axis named as `y_axis` and `x_axis` (optional).

Here is an example of a customized explicit encoding, where we want to check the variance of residuals. 

```{r std residual}
var_res_func = function(data) {
  data %>%
    mutate(y_axis = prediction - observation) %>%
    mutate(y_axis = sqrt(abs(y_axis / sd(y_axis))))
}

mpg_model %>% 
  mcplot() + 
  mc_condition_on(x = vars(mpg)) + 
  mc_layout_encoding(var_res_func) + 
  mc_model_lineribbon(alpha = .2) + 
  mc_model_point(n_sample = 1)
```


