---
title: "Specifying visual representations for model predictions and observed data using modelcheck"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing the distribution of model prediction and observed data using modelcheck}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette describes how to use the `modelcheck` package to specify the visual representations used to describe the distribution of model predictions and observed data. We developed `modelcheck` to support visual representations of three types: area/extent, visual variable, and countable icon. For a more general introduction to `modelcheck` and its use on a standard model check workflow, see `vignette("modelcheck")`.

# Setup

The following libraries are required to run this vignette:

```{r setup}
library(dplyr)
library(purrr)
library(modelcheck)
library(ggplot2)
library(ggdist)
library(cowplot)
library(rstan)
library(brms)
library(gganimate)

theme_set(theme_tidybayes() + panel_border())
```

These options help Stan run faster:

```{r}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Model

We use the built-in model `mpg_model`, a `brmsfit` object fitted in Gaussian model family with push-forward transformations `mu` and `sigma`, to demonstrate different visual representations `modelcheck` defines to show the distribution of model predictions and observed data. 


```{r example_model}
mpg_model
```

# Visualizing uncertainty using area/extent

`modelcheck` supports a list of visual representations to encode uncertainty using area/extent variable: 
- `mc_model_slab()` (`mc_obs_slab()`)
- `mc_model_ccdf()` (`mc_obs_ccdf()`)
- `mc_model_cdf()` (`mc_obs_cdf()`)
- `mc_model_eye()` (`mc_obs_eye()`)
- `mc_model_halfeye()` (`mc_obs_halfeye()`)
- `mc_model_histogram()` (`mc_obs_histogram()`)
- `mc_model_pointinterval()` (`mc_obs_pointinterval()`)
- `mc_model_interval()` (`mc_obs_interval()`)
- `mc_model_lineribbbon()` (`mc_obs_lineribbon()`)
- `mc_model_ribbon()` (`mc_obs_ribbon()`)

Examples:

```{r slab}
mpg_model %>%
  mcplot() +
  mc_model_slab(alpha = .5) +
  mc_obs_slab(alpha = .5) +
  mc_gglayer(coord_flip())
```

```{r slab x}
mpg_model %>%
  mcplot() +
  mc_model_slab(alpha = .5) +
  mc_obs_slab(alpha = .5) + 
  mc_condition_on(x = vars(vs)) +
  mc_gglayer(coord_flip())
```

```{r slab x interval}
mpg_model %>%
  mcplot() +
  mc_model_slab(alpha = .5) +
  mc_obs_interval(alpha = .5) + 
  mc_condition_on(x = vars(vs)) +
  mc_gglayer(coord_flip())
```

```{r lineribbon x pointinterval}
mpg_model %>%
  mcplot() +
  mc_model_lineribbon() +
  mc_obs_pointinterval() + 
  mc_condition_on(x = vars(disp))
```

# Visualizing uncertainty using visual variables

`modelcheck` supports a list of visual representation that encode uncertainty using diffferent visual variables: 
- `mc_model_gradientinterval()` (`mc_obs_gradientinterval()`)
- `mc_model_interval()` (`mc_obs_interval()`)
- `mc_model_lineribbbon()` (`mc_obs_lineribbon()`)
- `mc_model_ribbon()` (`mc_obs_ribbon()`)
- `mc_model_tile()` (`mc_obs_tile()`)

Examples:

```{r gradient x}
mpg_model %>%
  mcplot() +
  mc_model_gradientinterval() +
  mc_obs_point(shape = '|', size = 10) + 
  mc_condition_on(x = vars(vs)) +
  mc_gglayer(coord_flip())
```

```{r gradient x gradient}
mpg_model %>%
  mcplot() +
  mc_model_gradientinterval() +
  mc_obs_gradientinterval() + 
  mc_condition_on(x = vars(vs)) +
  mc_layout_nested() +
  mc_gglayer(coord_flip())
```

# Visualizing uncertainty using countable icons

`modelcheck` supports a list of visual representations that encode uncertainty using countable icons: 
- `mc_model_dots()` (`mc_obs_dots()`)
- `mc_model_dotsinterval()` (`mc_obs_dotsinterval()`)
- `mc_model_point()` (`mc_obs_point()`)
- `mc_model_line()` (`mc_obs_line()`)

Examples:

```{r dots x}
mpg_model %>%
  mcplot() +
  mc_model_dots(n_sample = 100) +
  mc_obs_dots() + 
  mc_gglayer(coord_flip())
```

```{r dots x raincloud}
mpg_model %>%
  mcplot() +
  mc_model_slab() +
  mc_obs_interval() +
  mc_obs_dots(side = "left") + 
  mc_condition_on(x = vars(vs)) +
  mc_gglayer(coord_flip())
```

# Customize the visual representation using modelcheck

If your visual representation is not supported directly by `modelcheck`, you can use `mc_model_custom()` and `mc_obs_custom()` to pass in a geom or stat function to specify the visual representation of model predictions and observed data. For example, we could use `mc_model_custom(stat_boxplot)` to specify a box plot for model predictions and `mc_obs_custom(geom_swarm)` to specify a swarm plot for observed data.

```{r customized plot}
mpg_model %>%
  mcplot() +
  mc_model_custom(stat_boxplot, notch = TRUE) +
  mc_obs_custom(geom_swarm) +
  mc_condition_on(x = vars(vs)) +
  mc_gglayer(coord_flip())
```
